\section{Method}
\label{method}

The methodology used is standard procedure in the domain of computational complexity. A brief review of this terminology is provided in~\nameref{sec:background}.
The process is as follows:

\begin{enumerate}
    \item It is proven that given a solution of a \textit{k-cleared-cells} instance, the solution is verifiable in polynomial time. By definition this will prove that \textit{k-cleared-cells} is in NP.
    \item A previously known NP-complete problem is reduced to \textit{k-cleared-cells}. Since a reduced problem can be no harder (in a computational sense) than the problem it is reduced to, it follows that \textit{k-cleared-cells} is NP-hard.
    \item If \textit{k-cleared-cells} is in NP and is NP-hard, it follows from definition that the \textit{k-cleared-cells} is NP-complete.
\end{enumerate}

\subsection{Finding a reduction}
As mentioned in~\autoref{subsub:sim} there exists many similarities between Lumines and the more thoroughly researched video game Tetris. This fact suggests that it is a good idea to draw inspiration from papers concerning the computational complexity of Tetris. In one such paper by \citeauthor{tetris}, the authors present a reduction from \textit{3-Partition} to the problem of how many rows can be cleared using some sequence of Tetris pieces \cite{tetris}. The implied notion of forcing game pieces to be placed in certain positions has showed valuable for the proof presented in this report. The other reports mentioned in \autoref{sec:background} has been helpful as well, giving valuable lessons on how to think about video game complexity. However, no properties as similar to Lumines as Tetris' properties has been found in the other games.

NP-hard problems priorly encountered in literature and courses were first examined when researching suitable problems to reduce from. Problems were considered on how good they could they could relate to the spatial characteristics of Lumines. Finally the \textit{Subset sum} problem was chosen, on the grounds that it was well understood by the authors and that mappings could be made intuitively from the problem to some characteristics of Lumines. For example it seemed plausible that a set of elements in a \textit{Subset sum} instance could be mapped to a set of Lumines blocks of in a sequence. The characteristic of summing elements in the \textit{Subset sum} problem also seemed like a good match for the stacking characteristic of Lumines gameplay.
